\documentclass[conference]{IEEEtran}


\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts,mathtools,amsthm}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{url}
\usepackage{tikz}
\usepackage{comment}


\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{claim}{Claim}
\newtheorem{lemma}{Lemma}
\newtheorem{fact}{Fact}
\newtheorem{corollary}{Corollary}
\newtheorem{rem}{Remark}
\newtheorem{observation}{Observation}
\newtheorem{definition}{Definition}
\newtheorem{property}{Property}




\DeclarePairedDelimiter\pair{(}{)}
\DeclarePairedDelimiter\set{\{}{\}}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

\newcommand\mahmoud[1]{\color{green}\textbf{\\ Mahmoud: #1}\\\color{black}}
\newcommand\stefan[1]{\color{blue}\textbf{\\ Stefan: #1}\color{black}}
\newcommand\maciek[1]{\color{brown}\textbf{\\ Maciek: #1}\color{black}}


\begin{document}

\section{Introduction}

Main technical contribution of this paper is $\Omega(k\ell)$ lower bound for any deterministic algorithm for Online Balanced Partition problem.
Previous known lower bound were $\Omega(k)$ (cite Marcin-Chen).
The gap between upper and lower bound is still substantial, as the best known algorithm is $O(k^2\ell^2)$ (cite Marcin-Chen).
We contribute to the search for the optimal algorithm by presenting $O(k\ell)$-competitive algorithm for the special case where the perfect partition of the input graph is assumed.

\maciek{Future directions of research: How well does the Perfect Partition Learner performs when applied to a single epoch in general online clustering problem (without the perfect partition?)? Epoch = C * l * (k-1) external requests.}

\section{The Model}

Formal introduction of the model.

 We say that an online algorithm DET is component-respecting if it never reaches a configuration with a component split between clusters.
  %Recall that if DET resides in a configuration with a component split between clusters, we issue intra-component requests until
 If DET splits a component, we issue intra-component requests until DET joins it.
 If it never joins a component, it is not competitive.
 
\input{lower}

\input{upper}

\appendix

\input{appendix}
	
\end{document}